the rethinking way of doing the binary outcome modeling

### 2021-03-15 scripts/54_rethinking_binary.Rmd
ok now it is time to move to the shotgun

    diff = abs( cols[[1]] -cols[[2]] )
    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
    return(cols)
  })
```

```{r}
# posterior prediction check
post <- extract.samples(mcr)
post_cr <- list(m = 1, p = 2) %>% 
  map(function(center_){
    cols = seq(1, length(div_value)) %>% 
      set_names(seq(1, length(div_value))) %>% 
      map(function(idx){
         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
      })

    diff = abs( cols[[1]] -cols[[2]] )
    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
    return(cols)
  }) 
```

```{r}
# the below is when the returned value is the difference between the low and high
# bind_rows(
#   bind_cols(prior_cr, grp ='prior', quant = c(0.025, 0.5, 0.975)),
#   bind_cols(post_cr, grp ='post', quant = c(0.025, 0.5, 0.975))
# )  %>% 
#   gather('center', 'value', m:p) %>% 
#   spread('quant', 'value') %>% 
#   mutate(grp = factor(grp, levels = c('prior','post'))) %>% 
#   ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
#   geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
#   ylim(0,1) 
```
```{r}
# plot the histogram of the probs corresponding to low and high diversity
# msk
tibble(
  prior_low = prior_cr %>% pluck('m') %>% pluck(1),
  prior_high = prior_cr %>% pluck('m') %>% pluck(2)
)%>% gather() %>% 
  gghistogram(x = 'value')+
  facet_grid(key ~ .)

@@

>>>>

    diff = abs( cols[[1]] -cols[[2]] )
    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
    return(ret)
  })
```
### posterior

```{r}
# posterior prediction check
post <- extract.samples(mcr)
post_cr <- list(m = 1, p = 2) %>% 
  map(function(center_){
    cols = seq(1, length(div_value)) %>% 
      set_names(seq(1, length(div_value))) %>% 
      map(function(idx){
         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
      }) 
    return(cols)
  }) 
```

```{r}
# plot the histogram of the probs corresponding to low and high diversity
# msk
tibble(
  prior_low = prior_cr %>% pluck('m') %>% pluck(1),
  prior_high = prior_cr %>% pluck('m') %>% pluck(2)
)%>% gather() %>% 
  gghistogram(x = 'value')+
  facet_grid(key ~ .)

@@

```{r}
# the difference of the probability 
post_cr <- list(m = 1, p = 2) %>% 
  map(function(center_){
    cols = seq(1, length(div_value)) %>% 
      set_names(seq(1, length(div_value))) %>% 
      map(function(idx){
         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
      })

```{r}
#the below is when the returned value is the difference between the low and high
diff_prob_post <- bind_rows(
  bind_cols(prior_cr, grp ='prior', quant = c(0.025, 0.5, 0.975)),
  bind_cols(post_cr, grp ='post', quant = c(0.025, 0.5, 0.975))
)  %>%
  gather('center', 'value', m:p) %>%
  spread('quant', 'value') %>%
  mutate(grp = factor(grp, levels = c('prior','post'))) %>%
  # only showing the post check value
  filter(grp == 'post')
# mean at both centers are 0.20 actually

diff_prob_post  %>% 
  ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 sd apart') +
  geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') 
```

```{r}
#  The two patients with highest and lowest diversity
# post only 
# lowest appear in one patient in p; highest is in one patient in M  
div_l <- min(meta$logdiv_s)
pl <- inv_logit( post$b * div_l + post$a[, 2] )

div_h <- max(meta$logdiv_s) 
ph <- inv_logit( post$b * div_h + post$a[, 1] )

bind_rows(
  tibble(value = pl, grp = 'lowest'),
  tibble(value = ph, grp = 'highest')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean', xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
              title = 'Predicted probability of CR d100 between the pt with highest div and the pt with lowest') 
```
